---
title: "Regression Model - Course Project"
author: "Jordan Woloschuk"
date: "9/21/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Motor Trend: Analysis on Variables and the Impact on Miles per Gallon

## 1) Executive Summary


The purpose of this analysis is to examine a collection of cars and determine the
relationship between vechicle variables (e.g. number of cyclinders) and fuel
economy (i.e. miles per gallon).

This analysis will seek to answer the following questions:

1. Is an automatic or manual transmission better for MPG?
2. Quantify the MPG difference between automatic and manual transmissions?


## 2) Data Processing

### 2.1) Loading Libraries

Load necessary libraries for data analysis and developing results.
```{r Libraries, warning=FALSE, message=FALSE}

library(ggplot2)
library(dplyr)
# GGally for ...........
library(GGally) 
# ggfortify
library(ggfortify)

```

### 2.2) Loading Data

Load necessary mtcars dataset.
```{r Data, warning=FALSE, message=FALSE}
data(mtcars)
```

### 2.3) Modifying Data 

The "am" variable will be converted to a factor, such that 0 = Automatic and 
1 = Manual transmission.Similarly,  varaible will also be converted to a factor. 

