---
title: "Regression Model - Course Project"
author: "Jordan Woloschuk"
date: "9/21/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Motor Trend: Analysis on Variables and the Impact on Miles per Gallon

## 1) Executive Summary


The purpose of this analysis is to examine a collection of cars and determine the
relationship between vechicle variables (e.g. number of cyclinders) and fuel
economy (i.e. miles per gallon).

This analysis will seek to answer the following questions:

1. Is an automatic or manual transmission better for MPG?
2. Quantify the MPG difference between automatic and manual transmissions?


## 2) Data Processing

### 2.1) Loading Libraries

Load necessary libraries for data analysis and developing results.
```{r Libraries, warning=FALSE, message=FALSE}

library(ggplot2)
library(dplyr)
# GGally for the ggpairs function
library(GGally) 
# ggfortify
library(ggfortify)

```

### 2.2) Loading Data

Load necessary mtcars dataset.
```{r Data, warning=FALSE, message=FALSE}
data(mtcars)
```

### 2.3) Modifying Data 

The variables "cyl" (# cyclinders), "vs" (engine shape), "am" (transmission),
"gear" (# of gears) will be converted to factor variables since they are not continous. 
```{r Factor_Data, warning=FALSE, message=FALSE}
data(mtcars)

mtcars_data <- mtcars # creating a new mtcars dataframe to be modified
mtcars_data$cyl <- as.factor(mtcars_data$cyl)

mtcars_data$vs <- as.factor(mtcars_data$vs)
# Set levels for the engine shape (V-shaped or Straight)
levels(mtcars_data$vs) <- c("V-shaped", "Straight")

mtcars_data$am <- as.factor(mtcars_data$am)
# Convert 0 = Automatic and 1 = Manual
levels(mtcars_data$am) <- c("Automatic", "Manual")

mtcars_data$gear <- as.factor(mtcars_data$gear)
```

## 3) Exploratory Data Analyses

```{r Expo_Data, warning=FALSE, message=FALSE}
head(mtcars_data,4) # Sample of the first 4 rows of data
dim(mtcars_data) # Dimensions of the mtcars_data
```

In order to understand the relationship between transmission type and fuel economy
we will develop a box plot to show the impact automatic/manual has on MPG.

Please refer to **Fig. 1** in the Appendix.

After review of **Fig. 1**, it can be seen that manual transmissions results in higher
MPG. However, it is not clear if other variables are also influencing this outcome,
and if manual transmission vechicles have higher MPG with all else equal. To do so
a pair graph will be developed. 






## TK) Appendix

### **Fig. 1**
```{r Fig_1, warning=FALSE, message=FALSE}
Fig_1 <- ggplot(mtcars_data, aes(am, mpg))
Fig_1 + geom_boxplot(aes(fill = am)) + xlab("Transmission")
```


